Air cargo load and route planning in pickup and delivery operations

### 2023-06-03 CAIE/results.tex
CAIE paper transferred to parallel folder

\begin{frontmatter}

\title{Air cargo load and route planning in pickup and delivery operations}

\author{A.C.P.~Mesquita}
\ead{celio@ita.br}

\author{C.A.A.~Sanches\corref{cor1}}
\ead{alonso@ita.br}
\cortext[cor1]{Corresponding author.}

\address {Instituto Tecnol\'{o}gico de Aeron\'{a}utica - DCTA/ITA/IEC\\
Pra\c{c}a Mal. Eduardo Gomes, 50\\
S\~{a}o Jos\'{e} dos Campos - SP - 12.228-900 - Brazil}

\begin{abstract}

In aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading, for fast take-off, and mission accomplishment, especially in times of crisis, disaster relief missions, short deadlines for contracting customers, or any external pressure for immediate take-off. {\color{blue} In addition, there are no commercially available systems that can assist the load and trip planners with the pallet building and aircraft loading and balancing demands for transport at each hub.} This enables other risks, such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We defined and solved the problem of planning the loading and routing of an aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs. This NP-hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs. We performed multiple experiments with a commercial solver and four well-known meta-heuristics on data based on the transport history of the {\it Brazilian Air Force}. We also designed a new heuristic that quickly finds good solutions for a wide range of problem sizes: an essential contribution, which solved all test scenarios within an acceptable operational run time.

\end{abstract}

\begin{keyword}
Load Planning \sep Air Palletization \sep Weight and Balance \sep Pickup and Delivery \sep Vehicle Routing
\end{keyword}

\end{frontmatter}

\label{sec1}
\section{Introduction}

Air cargo transport involves several sub-problems that are difficult to solve. Recently, Brandt and Nickel \cite[p. 401]{BrandtStefan2019} defined the {\it Air Cargo Load Planning Problem} (ACLPP) as four sub-problems: {\it Aircraft Configuration Problem} (ACP), {\it Build-up Scheduling Problem} (BSP), {\it Air Cargo Palletization Problem} (APP) and {\it Weight and Balance Problem} (WBP). Several aspects were considered in this modeling: characteristics of the items to be transported (dimensions, scores, dangerousness, etc.); types and quantities of {\it unit load devices} (ULDs), commonly called pallets; when these pallets are assembled; how items are allocated to pallets; in which positions these pallets are to be placed; how the total cargo weight is balanced; etc. They also presented a comprehensive bibliographic survey of solving methods that had been developed in different situations.

However, there are still other important challenges in air cargo transport that go beyond the definition of the ACLPP, especially with regard to the flight itinerary and the loading and unloading at each destination (or node) of this travel plan. In this context, at least two more important sub-problems can be considered: pickup and delivery operations at each node, called {\it Pickup and Delivery Problem} (PDP), and the search for the lowest cost route, which is the well-known {\it Traveling Salesman Problem} (TSP).

Considering air cargo transport, Table \ref{tab:sa} lists the main works in the literature and the corresponding sub-problems addressed. We also indicate whether the dimensions of the items were taken into account ({\bf 3D} or {\bf 2D}) and which solution method {\color{blue}was used: heuristic search methods ({\bf Heu}), integer programming ({\bf Int}), or linear programming ({\bf Lin}).}

\begin{table}[H]
	\centering
	\caption{Air cargo transport: literature, problems and features}  \label{tab:sa}
	\scriptsize
	\renewcommand{\arraystretch}{1.1} 
	\begin{tabular}{r|cccc|ccccc}
		\toprule
		 Work & {\bf APP}  & {\bf WBP}  &  {\bf PDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		Larsen and Mikkelsen \cite{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		Brosh \cite{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		Ng \cite{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		Heidelberg {\it et al.} \cite{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		Mongeau and Bes \cite{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		Fok and Chun \cite{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		Chan {\it et al.} \cite{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		Kaluzny and Shaw \cite{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		Verstichel {\it et al.} \cite{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		Mesquita and Cunha \cite{MesquitaCunha2011}   & $.$ & $.$    & $\bigstar$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\		
		Limbourg {\it et al.} \cite{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		{\color{blue}Roesener and Hall} \cite{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		Vancroonenburg {\it et al.} \cite{Vancroonenburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		Lurkin and Schyns \cite{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		Roesener and Barnes \cite{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		Paquay {\it et al} \cite{PaquaySchynsLimbourg2016,PaquayLimbourgSchynsOliveira2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		Chenguang {\it el al.} \cite{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		Wong and Ling \cite{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		Wong {\it et al.} \cite{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		Zhao {\it et al.} \cite{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		{\bf This article}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

As can be seen, so far Virginie Lurkin and Michaël Schyns \cite{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. Although it is innovative, strong simplifications were imposed by the authors: in relation to loading, APP was ignored; with regard to routing, it is assumed a pre-defined tour plan restricted to two legs. It is important to note that these authors consider an aircraft with two doors, and the minimization of loading and unloading costs at the intermediate node was modeled through a container sequencing problem. Referring directly to this work, Brandt and Nickel \cite[p. 409]{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes or the models omit some practically relevant constraints}.

There are real situations that are much more complex. In this work, we consider a practical case in Brazil, which is the largest economy in Latin America. Due to its dimensions, this country has the largest air market on the continent with $2,499$\/ registered airports, of which $1,911$\/ are private and $588$\/ are public. Although it is an immense distribution network, airlift missions consider 3 to 5 nodes per tour plan. Throughout this work, we address routes with up to 7 nodes, as can be seen in Table \ref{tab:costs} and Figure \ref{fig:nodes}.

{\color{blue} 
	This 7-node limitation is more than enough for the case of the {\it Brazilian Air Force}. In case some extension is needed for this range, a more suitable TSP solution method may be adopted, as full enumeration may make it impractical for other clients' real-world scenarios.
}

\begin{table}[H]

	\begin{minipage}{0.05\linewidth}

	\end{minipage}\hfill 
	\begin{minipage}{0.45\linewidth}

		\caption{Brazilian airports distances ($km$)}  \label{tab:costs}
		\centering
		\footnotesize
		\newcolumntype{Y}{>{\centering\arraybackslash}p{0.09\textwidth}}
		\newcolumntype{X}{>{\centering}p{0.09\textwidth}}

		\begin{tabular}{X X X X X X X Y}
			\toprule
			Node & $l_0$ & $l_1$ & $l_2$ & $l_3$ & $l_4$ & $l_5$ & $l_6$ \\
			IATA*   & GRU   & GIG   & SSA   & CNF   & CWB   & BSB   & REC \\	
			\midrule	
			GRU     & 0	    &343	&1,439   &504    &358    &866    &2,114\\
			GIG	    & 343	&0	    &1,218   &371    &677    &935    &1,876\\
			SSA	    & 1,439	&1,218	&0	    &938    &1,788   &1,062   &676\\
			CNF	    & 504	&371	&938	&0	    &851    &606    &1,613\\
			CWB	    & 358	&677	&1,788	&851	&0	    &1,084   &2,462\\
			BSB	    & 866	&935	&1,062	&606	&1,084	&0	    &1,658\\
			REC	    & 2,114	&1,876	&676	&1,613	&2,462	&1,658	&0\\
			\bottomrule
			\multicolumn{8}{c}{*International Air Transport Association}\\
			\multicolumn{8}{c}{\small\textsuperscript{Source: www.airportdistancecalculator.com}}\\
		\end{tabular}
		\normalsize

	\end{minipage}\hfill 
	\begin{minipage}{0.50\linewidth}
		\centering
		\includegraphics[scale=0.25]{Images/nodes.png}
		\captionof{figure}{A route with some Brazilian airports}
		\label{fig:nodes}		
	\end{minipage}
\end{table}

@@

	Although there are many other airports in the Brazilian Air Force's interest, these seven were chosen due to their huge demands and short lead times for transportation, which are worthy of aerial transportation. The other airports may have a lower transportation demand, which is frequently less costly to address by cabotage, railways, or road transport as well.
}

In the {\it Brazilian Air Force}\/ missions, hundreds of items can be carried at each node, where the objectives are to prioritize the transport of the most important items and minimize the cost of fuel along the route. As standardized pallets are used with predefined positions on the aircraft, it is possible to carry out loading and unloading at each node in around 30 minutes{\color{blue}, if the pallet's position and contents are already planned}.

{\color{blue}
As to the cargo item arrangement on the pallets as well as the route planning, which normally takes two hours, which is the same time reported by Kelly Fok and Andy Chun \cite{fok2004optimizing} "the container load planning is usually done in roughly 2 hours before departure, when all cargo details are in place.", it is essential to find a way to shorten this time to improve the efficiency of cargo transportation. However, there are no commercially available systems or software packages that can simultaneously solve these problems.
}

\begin{table}[H]

	\begin{minipage}{0.05\linewidth}

	\end{minipage}\hfill 
	\begin{minipage}{0.40\linewidth}

	{\color{blue}	
Cargo handling at modern airports usually involves powerful equipment like dollies, powered conveyors, elevating transfer vehicles, caster decking, scissor lifts, electrical controls, hydraulically adjustable heights, forklift trucks, and plenty of room in the airport cargo terminal.

\vspace{3mm}
These facilities can help with a complete pallet re-arrangement in the cargo bay, if necessary, in less than 30 minutes if the pallet position contents are already planned.
}

	\end{minipage}\hfill 
	\begin{minipage}{0.55\linewidth}
		\centering
		\includegraphics[scale=0.2]{Images/handling.png}
		\captionof{figure}{Cargo handling equipment}
		\label{fig:handling}

		\small\textsuperscript{Source: NARA \& DVIDS Public Domain Archive}	
	\end{minipage}
\end{table}

This work proposes a method that {\color{blue}addresses} these objectives: a pallet building and arrangement plan with a routing option that maximizes the benefit-cost ratio for the smooth execution of pickup and delivery transport missions.

We develop a heuristic that can be executed on a simple handheld computer (like a laptop or a tablet) and that provides a solution quickly enough to keep this cargo handling time {\color{blue}under 30 minutes.} This heuristic {\color{blue} solution certainly will} reduce the stress that transport planners are subjected to, because they have to deal with a lot of information in planning the aircraft route, assembling the pallets ({\color{blue}regarding their positions), and a pick up and delivery plan to each node.}

To the best of our knowledge, this is the first time that an air cargo transport problem that simultaneously involves APP, WBP, PDP and TSP has been addressed. This new problem is named {\it Air Cargo Load Planning with Routing, Pickup and Delivery Problem} (ACLP+RPDP).

This article is organized into six more sections. In Section \ref{sec2}, we give a brief review of the literature. In Section \ref{sec3}, we present the problem context and assumptions, and in Section \ref{sec4}, we describe the mathematical model and how we dealt with its issues. In Section \ref{sec5}, we describe the elaborate algorithms, whose results are presented in Section \ref{sec6}. Finally, our conclusions are in Section \ref{sec7}.

\section{Related literature}
\label{sec2}

In this section, we briefly describe the characteristics of the main works related to air cargo transport, following the chronological order of Table \ref{tab:sa}.

Ole Larsen and Gert Mikkelsen \cite{LarsenMikkelsen1979} developed an interactive procedure for loading 14 types of Boeing 747 into a two-leg transportation plan. Seven types of items were considered to be allocated in 17 to 42 positions. With non-linear programming and heuristics, they present a solution that minimizes positioning changes in the intermediate node, optimizing the load balancing in the aircraft.

Israel Brosh \cite{Brosh1981} addressed the problem of planning the allocation of cargo on an aircraft. Considering volume, weight and structural constraints, the author finds the optimal load layout through a fractional programming problem.

Kevin Ng \cite{Kevin1992} developed a multi-criteria optimization approach to load the {\it C-130} aircraft of the {\it Canadian Air Force}. Based on integer programming, this model provides timely planning and improves airlift support for combat operations, solving WBP with pallets in fixed positions, and considering 20 different items.

Heidelberg {\it et al.} \cite{Heidelberg1998} developed a heuristic for 2D packing in air loading, comparing it with methods for solving the {\it Bin Packing Problem}. These authors conclude that the classical algorithms are inadequate in this context, because they ignore the aircraft balancing constraints.

Marcel Mongeau and Christian Bes \cite{MongeauBes2003} presented a method based on linear integer programming to solve the problem of choosing and positioning containers in the {\it Airbus 340-300}. Safety and stability constraints were considered, with the objective of minimizing fuel consumption.

Kelly Fok and Andy Chun \cite{fok2004optimizing} developed a web-based application to make efficient use of space and load balancing for an air cargo company. Based on an analysis of historical data, an operational load planning with mathematical optimization is obtained. This container load planning is usually done in roughly 2 hours before departure, when all cargo details are in place.

Chan {\it et al.} \cite{Chan2006} carried out a case study with heterogeneous pallets. To minimize the total cost of shipping, they developed a 3D packing heuristic, with a loading plan for each pallet. Although the authors do not consider load balancing or positioning of pallets in the cargo hold, this method is relevant in commercial and industrial applications, where cargo items tend to be less dense.

Bohdan Kaluzny and David Shaw \cite{KaluznyBohdanL2009Oalb} developed a mixed integer linear programming model to arrange a set of items in a military context that optimizes the load balance. {\color{blue}The objective function can be chosen to minimize the deviation of the center of gravity from the target position or to maximize the function of the items loaded. This approach does not palletize items, which are arranged in the cargo bay by a 2D packing procedure.
}

Verstichel {\it et al.} \cite{Verstichel2011} solved WBP by selecting the most profitable subset of containers to be loaded into an aircraft by the use of mixed-integer programming. Experimental results on real-life data showed significant improvements compared to those obtained manually by an experienced planner.

Mesquita and Cunha \cite{MesquitaCunha2011} presented a heuristic for a real problem of the {\it Brazilian Air Force}, which consists of defining transport routes with simultaneous collection and delivery from a central distribution terminal.

Limbourg {\it et al.} \cite{Limbourg2012} developed a mixed-integer program for optimally rearranging a set of pallets into a compartmentalized cargo aircraft, specifically the {\it Boeing 747}.
{\color{blue}These authors approach regarding CG deviation is different from ours because we simply position the CG within its permitted range with the sole objective of maximizing the transportation value. As the longitudinal stability is handled by fly-by-wire-equipped aircraft with automatic longitudinal trimming, it is expected to increase our proposed method's ability to offer a solution to carry additional freight.
}

August Roesener and Shane Hall \cite{RoesenerHall2014} solved APP and WBP as an integer programming problem {\color{blue} where the items are selected for shipment based on a utility score, then assigned to pallets, which will be loaded into an aircraft in a specific pallet position.
The pallets are then packed in a manner to optimize both the pallet and aircraft characteristics, such as item utility, pallet occupancy, and aircraft center of balance.
This paper groups items by destinations to be served by multiple aircraft with multiple destinations.

Differently from these authors' work, we plan to meet all transportation demands in a set of nodes with one aircraft in a simultaneous delivery and pickup tour.
}

Vancroonenburg {\it et al} \cite{Vancroonenburg2014} presented a mixed integer linear programming model that selects the most profitable pallets, satisfying safety and load balancing constraints on the {\it Boeing 747-400}. Using a solver, these authors solved real problems in less than an hour.

As already mentioned, Virginie Lurkin and Michaël Schyns \cite{LurkinSchyns2015} was the first work that simultaneously modeled WBP and PDP in air cargo transport. The authors demonstrated that this problem is NP-hard and performed some experiments with real data, noting that their model offers better results than those obtained manually.

Roesener and Barnes \cite{RoesenerBarnes2016} proposed a heuristic to solve the {\it Dynamic Airlift Loading Problem} (DALP). Given a set of palletized cargo items that require transport between two nodes in a time frame, the objective of this problem is to select an efficient subset of aircraft, partition the pallets into aircraft loads and assign them to allowable positions in those aircraft.

Paquay {\it et al.} \cite{PaquaySchynsLimbourg2016} presented a mathematical modeling to optimize the loading of heterogeneous 3D boxes on pallets with a truncated parallelepipeds format. Its objective is to maximize the volume used in containers, considering load balancing constraints, the presence of fragile items and the possibility of rotating these boxes. Paquay {\it et al.} \cite{PaquayLimbourgSchynsOliveira2018} developed some heuristics to solve this problem. {\color{blue} We also plan to apply these authors' technique to the ACLP part of our problem if we manage to keep the running time within one hour.}

Chenguang {\it et al.} \cite{YangLiuGao2018} modeled the air transport problem as a 2D packing problem, and presented a heuristic for its optimization in several aircraft, considering load balancing to minimize fuel consumption.
{\color{blue}We also plan to select the best tour that minimizes fuel consumption through cargo balancing, as it is known (see Mongeau and Bes \cite{MongeauBes2003} for further details) that cargo unbalancing may lead to extra fuel consumption.}

Wong and Ling \cite{wong2020}A developed a mathematical model and a tool based on mixed integer programming for optimizing cargo in aircraft with different pallet configurations. Balance constraints and the presence of dangerous items were considered. Wong {\it et al.} \cite{eugene2021} integrated this tool to a digital simulation model, with a visualization and validation system, based on sensors that alert about load deviations.

Zhao {\it et al.} \cite{zhao2021} proposed a new modeling for WBP based on mixed integer programming. Instead of focusing on the center of gravity (CG) deviation, the authors consider the original CG envelope of the aircraft, with a linearization method for its non-linear constraints.

As can be seen, none of these works address air cargo palletization and load balancing with route optimization in a multi-leg transportation plan. This is the objective of our work: to model and elaborate heuristics for a real problem that simultaneously involves 4 intractable sub-problems: APP, WBP, PDP and TSP.

\section{Problem context and assumptions}
\label{sec3}

In this section, we describe the context of the problem addressed in this work, as well as the assumptions considered.

\subsection{Operational premises}

As we are dealing with an extremely complex and diverse problem, we decided to establish some simplifying characteristics:

\begin{itemize}

	\item At each node of the tour, the items to be allocated are characterized by weight, volume {\color{blue}(the product of their dimensions, as most of them have a parallelepiped shape)}, scores, and previously known destinations. {\color{blue}But, as the dimensions are unknown, we leave the consideration of 3D packing to the next step of this work if we manage to keep the worst running time within one hour.}

	\item We considered a unique pallet type: the {\it 463L Master Pallet}, a common size platform for bundling and moving air cargo. It is the primary air cargo pallet for more than 70 Air Forces and many air transport companies. This pallet has a capacity of $4,500 kg$ and $13.7 m^3$ {\color{blue}(which may be limited by its position along the cargo bay)}, is equipped for locking into cargo aircraft rail systems, and includes tie-down rings to secure nets and cargo loads, which in total weighs $140 kg$. For more information, see {\tt www.463LPallet.com}.

	\item All items allocated on a pallet must have the same destination. A pallet which has not yet reached its destination may receive more items, although it is known that these operations of removing restraining nets increase handling time and the risk of improper delivery. We do not consider oversized cargo in this work, but only {\color{blue}cargo items} that fit on these pallets.

	\item Finally, as we are interested in minimizing fuel costs, {\color{blue}we disregarded some fuel costs not directly associated with aircraft flight, such as handling.}

\end{itemize}

Throughout this text, we call a {\color{blue} {\it Packed Contents (PC)} } a set of items of the same destination stacked on a pallet and covered with a restraining net. It is considered unique, having the same attributes of its components, whose values are the sum of individual scores, weights and volumes. See Figure \ref{fig:larger2}. To ensure accuracy in pickup and delivery operations, a PC must remain on board until its destination.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{Images/consolidated.jpg}
	\caption{Packed Contents on a 463L pallet inside a {\it Boeing C-17}}
	\small\textsuperscript{Source: From Wikimedia Commons, the free media repository}
	\label{fig:larger2}
\end{figure}

\subsection{Aircraft and load balancing}

We consider real scenarios with an aircraft with payload of 75,000 kg (with enough fuel for a range of 4,400km). The aircraft layout is presented in Figure \ref{fig:larger}, where the pallets positions are identified by $p_i$.

\begin{figure}[!h]
	\centering

	\begin{tikzpicture}[scale=1.2, samples=100]

		\filldraw[fill=blue!5!white, draw=black] (0, 0) rectangle (10.23, 2.6);

		\filldraw[fill=gray!20!white, draw=black] (0.20, 1.40) rectangle node{$p_{17}$} (1.43, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (0.20, 0.20) rectangle node{$p_{18}$} (1.43, 1.20);

		\filldraw[fill=gray!20!white, draw=black] (1.63, 1.40) rectangle node{$p_{15}$} (2.86, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (1.63, 0.20) rectangle node{$p_{16}$} (2.86, 1.20);

		\filldraw[fill=gray!20!white, draw=black] (3.06, 1.40) rectangle node{$p_{13}$} (4.29, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (3.06, 0.20) rectangle node{$p_{14}$} (4.29, 1.20);

		\filldraw[fill=gray!20!white, draw=black] (4.49, 1.40) rectangle node{$p_{11}$} (5.72, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (4.49, 0.20) rectangle node{$p_{12}$} (5.72, 1.20);

		\filldraw[fill=gray!20!white, draw=black] (5.92, 1.40) rectangle node{$p_{9}$} (7.15, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (5.92, 0.20) rectangle node{$p_{10}$} (7.15, 1.20);

		\filldraw[fill=gray!20!white, draw=black] (7.35, 1.40) rectangle node{$p_{7}$} (8.58, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (7.35, 0.20) rectangle node{$p_{8}$} (8.58, 1.20);

		\filldraw[fill=gray!20!white, draw=black] (8.78, 1.40) rectangle node{$p_{5}$} (10.1, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (8.78, 0.20) rectangle node{$p_{6}$} (10.1, 1.20);

		\filldraw[fill=blue!5!white, draw=black] (10.23, 0) rectangle (13.27, 2.6);
		\filldraw[fill=gray!20!white, draw=black] (10.41, 1.40) rectangle node{$p_{3}$} (11.64, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (10.41, 0.20) rectangle node{$p_{4}$} (11.64, 1.20);
		\filldraw[fill=gray!20!white, draw=black] (11.84, 1.40) rectangle node{$p_{1}$} (13.07, 2.40);
		\filldraw[fill=gray!20!white, draw=black] (11.84, 0.20) rectangle node{$p_{2}$} (13.07, 1.20);

	\end{tikzpicture}

	\begin{tikzpicture}[scale=1.2, samples=100]
		\draw[black, thick][<-]  (1, 2.77) node[anchor=east, font=\fontsize{7}{3.5}\selectfont]{Forward} -- (9, 2.77) node[anchor=west, font=\fontsize{7}{3.5}\selectfont]{Ramp} ;
	\end{tikzpicture}

	\caption{Aircraft layout} \label{fig:larger}
\end{figure}

The torque applied to the aircraft must keep its CG in the operational range, which corresponds to a fixed percentage of the {\it Mean Aerodynamic Chord} \footnote{Chord is the distance between the leading and trailing edges of the wing, measured parallel to the normal airflow over the wing. The average length of the chord is known as the {\it Mean Aerodynamic Chord} (MAC).} which is considered $1.17m$ for the aircraft of this work. See Figure \ref{fig:lateral}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.22]{Images/lateral.png}
	\caption{Aircraft longitudinal cut showing red lines as pallets positions}
	\label{fig:lateral}
\end{figure}

Table \ref{tab:larger} shows the parameters. {\it $D_i^{long}$} and {\it $D_i^{lat}$} refer to the relative distances of pallet centroids (in meters) in relation to the CG of aircraft along both axes.

%Note that since the smaller aircraft has a single row of pallets positions, the y-axis passes through the geometric center of each pallet position.  Since the value is 0 for each position, $D_i^{lat}$ is not provided for the smaller aircraft.

Throughout this work, when we refer to the word "torque", we mean item or PC weight x pallet centroid distance to the CG ($D_i^{long}$ or $D_i^{lat}$), considering the angle between these dimensions to always be 90°. In both aircraft, as the ramps have an inclination of 25°, we made the necessary corrections in $D_i^{long}$, {\it Weight} and {\it Volume limits} of the corresponding pallets positions (see Table \ref{tab:larger}).

\begin{table}[H]
	\centering
	\caption{Aircraft parameters}  \label{tab:larger}
	\footnotesize
	\begin{tabular}{c | c c c c c c c c c}
		\toprule
		\textbf{Limits}& \multicolumn{3}{c}{$Payload$: 75,000kg} & \multicolumn{3}{c}{$limit^{CG}_{long}$: $1.170m$} &
		\multicolumn{3}{c}{$limit^{CG}_{lat}$: $0.19m$} \\
		\midrule
		\multirow{2}{*}{{\bf Pallets} ($p_i$)}  & $p_{17}$ & $p_{15}$ & $p_{13}$ & $p_{11}$ & $p_{9}$ & $p_{7}$ & $p_{5}$ & $p_{3}$ & $p_{1}$ \\
		& $p_{18}$ & $p_{16}$ & $p_{14}$ & $p_{12}$ & $p_{10}$ & $p_{8}$ & $p_{6}$ & $p_{4}$ & $p_{2}$ \\
		\midrule 
		\multirow{2}{*}{\boldm{$D_i^{long}$} ($m$)} & -17.57 & -13.17 & -8.77 & -4.40 & 0 & 4.40 & 8.77 & 11.47 & 14.89 \\
		& -17.57 & -13.17 & -8.77 & -4.40 & 0 & 4.40 & 8.77 & 11.47 & 14.89 \\			
		\midrule 
		\multirow{2}{*}{\boldm{$D_i^{lat}$} ($m$)}  & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 & 1.32 \\
		& -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 & -1.32 \\	
		\midrule
		{\bf Weight limits} ($W_i,\ kg$)      &   4,500   &    4,500  &   4,500   &  4,500    & 4,500     & 4,500     & 4,500     & 4,000    & 3,000   \\
		{\bf Volume limits} ($m^3$)   &   14.8   &   14.8   &  14.8    &  14.8    & 14.8     & 14.8     & 14.8     & 10.0    & 7.0 \\	
		\midrule

		\textbf{Fuel cost rate}  & \multicolumn{9}{c}{ $c_d$ = US\$ $4.900/km$ }	 \\

		\midrule

		\textbf{Fuel consumption limit rate}  & \multicolumn{9}{c}{ $c_g$ = $5\%$} \\

		\midrule

		\textbf{Maximum weight capacity}  & \multicolumn{9}{c}{ $W_{max}$ = $\sum_{i=1}^m W_i$} \\

		\bottomrule
	\end{tabular}
	\normalsize 
\end{table}

The {\it Fuel consumption limit rate} in Table \ref{tab:larger} is the percentage limit of cost increase  due to the CG deviation on the longitudinal axis of the aircraft (this is an arbitrated value for this work that should be better evaluated in future research, as it depends on the characteristics of the aircraft.).

It is important to consider that the range $[0,\ c_g]$ tends to zero as the aircraft attitude tends to level. As the CG deviation varies from 0 to the longitudinal CG limit, the leg cost increase rate varies from 0 to $c_g$.
}

\vspace{3mm}
We also make the following assumptions:
\begin{itemize}
	\item on each pallet, the items are distributed in such a way that their CG coincides with the centroid of the pallet;
	\item the CG of the payload must be at a maximum longitudinal distance of $limit^{CG}_{long}$ from the CG of the aircraft;
	\item the CG of the payload must be at a maximum lateral distance of $limit^{CG}_{lat}$ from the CG of the aircraft;
	\item pallets positions are distributed in two identical rows (with odd and even indices, respectively), and their centroids are at a distance $D^{lat}_{i}$ (where $i$ is the pallet index) from the center-line of the aircraft.
\end{itemize}

\subsection{Problem Summary}

\newcolumntype{M}{>{\raggedright}p{0.03\textwidth}}
\newcolumntype{T}{>{\raggedright\arraybackslash}p{0.9\textwidth}}

\bgroup
\def\arraystretch{1.2}
\begin{table}[H]
	\centering
	\small
	\begin{tabular}{MT}
		&Informally, ACLP+RPDP can be summarised as follows:\\
		\midrule
		max &  (items score sum) / (tour cost) of picked up and delivered items at each node on a tour  \\
		\midrule
s.t.    & Along a tour, the set of unvisited nodes is updated. \\
		& In each node, an item may be included in at most one pallet.\\		
		& In each node, {\it Packed Contents (PC)} are composed of items with the same destination. \\
		& Weight, volume and score of a PC are the corresponding sum of its components.\\
		& A PC remains on board until its destination.\\	
		& Items and a PC can only be included in the same pallet if their destinations are the same.\\
		& Only items destined to the remaining nodes can be loaded.  \\
		& The lateral and longitudinal torques must be within the operational range of the aircraft.\\
		& Weight and volume limitations of the pallets must be respected.\\
		& The total weight must be less than the aircraft payload or the total pallet capacity, whichever is the lowest.\\	
		\midrule
	\end{tabular}
	\normalsize
\end{table}
\egroup

\section{The mathematical modeling}
\label{sec4}

Given the assumptions, scenarios and parameters described in the previous section, we are ready to present the mathematical modeling of the ACLP+RPDP.

\vspace{3mm}
\textbf{Problem structure and input data:}

Let $C=\{c_{ab}\}$ be the cost matrix of flights, where $c_{ab} = c_d*d(l_a,l_b), 0 \leq a,b \leq \mbox{K}$.

Let $S_K = \{s: \{1, \dots, \mbox{K}\} \rightarrow \{1, \dots, \mbox{K}\} \}$ be the set of $\mbox{K}!$ permutations indexed by $\pi$, which correspond to all possible tours (or itineraries) that have node $0$ as origin and end, passing through the other $\mbox{K}$ nodes.

Note that $\pi_k$ is the node in position $k$ of tour $\pi$, and $\pi_0$ is always node 0, the base.

$\tau^{\pi_k}$ is the total torque applied by the loaded pallets to the aircraft in this node.

Let $L = \{ 0, \pi_1, \ldots, \pi_{K} \}$ be a tour with $\mbox{K}+1$ nodes.

Let $d(\pi_a,\ \pi_b)$ be the distance from $\pi_a$ to $\pi_b$, where $0 \leq a,\ b \leq \mbox{K}$. By definition, $d(\pi_a,\ \pi_a)\ =\ 0$.

Let $L^{\pi_k}$ be the set of remaining nodes when the aircraft is in $\pi_k$, $0 \leq k \leq \mbox{K}$. Therefore, $L^{\pi_0}\ =\ L$ and $L^{\pi_K}\ =\ \{ 0 \}$.

Let $M^{\pi_k} = \{1, 2, \ldots, m \}$ the set of $m$ empty pallets assigned to specific positions within an aircraft. Each pallet $i$, has weight capacity $W_i$, volume capacity $V_i$, pallet destinations $to^{\pi_k}_i$, $0 \leq k \leq K$, the longitudinal distance to the CG of aircraft $D^{long}_i$, and the lateral distance to the center line of aircraft $D^{lat}_i$. Note that all pallets and aircraft attributes are in uppercase.

>>>>

@@

\subsection{Results obtained}
In the tests performed, we used a 64-bit, 16GB, 3.6GHz, eight-core processor with {\it Linux Ubuntu 22.04.1 LTS 64-bit} as the operational system and {\it Python 3.10.4}\/ as the programming language. We also used the well-known solver {\it Gurobi}\/ (www.gurobi.com), version 9.5.2.
